Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-11 article/article.tex
Preprocessing draft

Here we will need a nice picture, perhaps a pagewide one with labels of data for 1 trial?

\subsection{Preprocessing}
\label{ssec:preprocessing}

@@

For each classifier, a ``leave one (subject) out" cross validation procedure was used to avoid overfitting. For $N=6$ iterations, $N-1$ subjects functioned as training set, while the remaining subject served as validation set. The classifier accuracy is the average accuracy over all iterations. 
We compared the following classifiers:\\

\textbf{\emph{Gaussian Naive Bayes (GNB).}} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}. It assumes a Gaussian distribution (cf.~Eq.~\ref{eq:gaussianDistribution}) and independent input features.

@@

where $C_i$ denotes the probability of the class---sentence or picture. Two approaches were used to estimate the parameters of the model. In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan separately. In both methods the decision of the classifier is given by Eq.~\ref{eq:maxLikelihood}, which is the maximum likelihood of the two classes over all scans.

\begin{equation}
\label{eq:maxLikelihood}
\operatorname{arg\,max}_i \sum_{t=1}^{N}{\log{P(C_{i}|X)}}
\end{equation}

where $t$ denotes the scan index. The advantage of the first method is that fewer parameters need to be estimated with greater amount of examples. It thus reduces the space and time complexity, and variance of the model. However, because it is likely that the assumptions of equal means and standard deviations for each scan do not hold, the second approach may decrease the inductive bias and enhance the accuracy.

\textbf{\emph{Linear Support Vector Machine (LSVM).}} The LSVM is a discriminant based approach to classification. For an elaborate description see \cite{bk:2004aplay}. For each selected scan belonging to a stimulus, the LSVM classifies it as either a picture or a sentence. The final classification of the stimulus is

\textbf{\emph{Multivariate Gaussian (MVG).}} In MVG, independence between feature vectors is not assumed. Instead, the mutivariate distribution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters. Again, the increase in parameters to be estimated compared to GNB (the correlations) may decrease generality, while also decreasing inductive bias.

>>>>

\subsection{Preprocessing}
\label{ssec:preprocessing}
The dataset contains activation values for each voxels with respect to a baseline activation value for a subject (measured before performing the task). To make optimal use of the information value of each voxel, we have created normalized copies of the dataset to use as input for the algorithms. The effect of this preprocessing step depends on assumptions made about the form of our data, which are not available. Instead, we had to verify the effect by looking at the resulting classification results (see table~\ref{results}). As a first step, we applied row normalization by setting the mean activation value to zero for each trial. The assumption was that the relative activation has more predictive power than the actual value. We then continued by transforming these values to z-scores (by setting the standard-deviation to 1) but this turned out to negatively impact accuracy (indicating that the transformation lost information). For completeness, we also performed normalization of higher-order properties (such as skew and kurtosis) but found a significant further decrease in accuracy. We therefore did not further researched performance on these sets. One other approach we tried was the application of column normalization, setting the mean value across trials to zero.

@@

For each classifier, a ``leave one (subject) out" cross validation procedure was used to avoid overfitting. For $N=6$ iterations, $N-1$ subjects functioned as training set, while the remaining subject served as validation set. The classifier accuracy is the average accuracy over all iterations. 
We compared the following classifiers:

\emph{Gaussian Naive Bayes (GNB).} This classifier builds a generative model underlying a class, by estimating the means and the standard deviations of the input features from the training set \cite{bk:2004aplay}. It assumes a Gaussian distribution (cf.~Eq.~\ref{eq:gaussianDistribution}) and independent input features.

@@

where $C_i$ denotes the probability of the class---sentence or picture. Two approaches were used to estimate the parameters of the model. In the first method, averages of the selected scans belonging to a stimulus were used to estimate the parameters, while in the second method, the parameters were estimated for each scan separately. In both methods, the resulting log-probability of a class was the sum of log-likelihoods per scan and the class with maximum log-probability was selected. The advantage of the first method is that fewer parameters need to be estimated with greater amount of examples. It thus reduces the space and time complexity, and variance of the model. However, because it is likely that the assumptions of equal means and standard deviations for each scan do not hold, the second approach may decrease the inductive bias and enhance the accuracy.

Linear Support Vector Machine (LSVM). The LSVM is a discriminant based approach to classification. For an elaborate description see \cite{bk:2004aplay}. For each selected scan belonging to a stimulus, the LSVM classifies it as either a picture or a sentence. The final classification of the stimulus is

Multivariate Gaussian (MVG). In MVG, independence between feature vectors is not assumed. Instead, the mutivariate distribution is used to estimate the probability of the classes, with the covariance matrix and the means as parameters. Again, the increase in parameters to be estimated compared to GNB (the correlations) may decrease generality, while also decreasing inductive bias.
